---
title: "explore likelihood"
author: "Victor Møller Poulsen"
output:
  pdf_document: default
  html_document: default
---

```{r, include=FALSE}

# consider pacman
install.packages("pacman") 
library(pacman)

p_load(tidyverse, brms, ggthemes, bayesplot, cowplot, tidybayes, modelr)

# set up cmdstanr 
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
library(cmdstanr)
install_cmdstan(cores = 2, overwrite = TRUE)

```

# load data 

```{r}

# check data
d <- read_csv("/work/50114/MAG/data/modeling/psych_replication_matched.csv") %>%
  mutate(log_teamsize = log(n_authors), 
         condition_coded = ifelse(condition == "experiment", 1, 0),
         condition_fct = as_factor(condition), 
         teamsize_scaled = (n_authors-min(n_authors))/(max(n_authors)-min(n_authors)),
         days_after_2010_scaled = days_after_2010/max(days_after_2010),
         id_fct = as_factor(PaperId)) %>% # because min = 0
  glimpse()


```

# read models (from fit_likelihood.Rmd)

```{r}

zip_prior <- readRDS("/work/50114/MAG/modeling/models/zip_prior.rds")
zip_post <- readRDS("/work/50114/MAG/modeling/models/zip_post.rds")
negbin_prior <- readRDS("/work/50114/MAG/modeling/models/negbin_prior.rds")
negbin_post <- readRDS("/work/50114/MAG/modeling/models/negbin_post.rds")
zinegbin_prior <- readRDS("/work/50114/MAG/modeling/models/zinegbin_prior.rds")
zinegbin_post <- readRDS("/work/50114/MAG/modeling/models/zinegbin_post.rds")

```

# prior predictive checks 

```{r}

prior_check <- function(model, ndraws, title, xmax){
  
  pp_check(model, 
           ndraws = ndraws) +
    labs(title = title) + 
    theme_minimal() + 
    xlim(0, xmax)
  
} 

```

## ZIP prior

```{r}

prior_check(zip_prior, 100, "zip prior (x cutoff: 25)", 25)

```

## negative binomial prior

```{r}

prior_check(negbin_prior, 100, "negative binomial prior (x cutoff: 25)", 25)

```

## zero-inflated negative binomial prior 

```{r}
prior_check(zinegbin_prior, 100, "zero-inflated negative binomial prior (x cutoff: 25)", 25)
```

# plot traces 
reasonable for negative binomial and for zero-inflated negative binomial 

```{r}

plot(zip_post, N = 3) # gives weird stuff 
plot(negbin_post, N = 3) # much more reasonable
plot(zinegbin_post, N = 3) # also reasonable 

```

# check posterior sampling 
https://mc-stan.org/bayesplot/articles/graphical-ppcs.html

```{r}

y <- d$c_5
y_zip <- posterior_predict(zip_post, draws = 500)
y_negbin <- posterior_predict(negbin_post, draws = 500)
y_zinegbin <- posterior_predict(zinegbin_post, draws = 500)

```

# overlay & histogram

## ZIP
we see the key issue again here... 
we have two distributions which is not reflective of the actual data. 

```{r}

ppc_dens_overlay(y, y_zip[1:50, ]) + xlim(0, 50)
ppc_hist(y, y_zip[1:5, ]) + xlim(-1, 50)


```

## negative binomial 
looks really good. 
has a hard time figuring out the correct number of 0s. 

```{r}

ppc_dens_overlay(y, y_negbin[1:50, ]) + xlim(01, 50)
ppc_hist(y, y_negbin[1:5, ]) + xlim(-1, 50)


```

## zero-inflated negative binomial 
looks really good. 
very similar to negbin.

```{r}

ppc_dens_overlay(y, y_zinegbin[1:50, ]) + xlim(-1, 50)
ppc_hist(y, y_zinegbin[1:5, ]) + xlim(-1, 50)

```

# more PPC 

## ability of models to produce zeros 
All models pretty good at producing the right amount of zeros. 

```{r}

prop_zero <- function(x) mean(x == 0)
ppc_stat(y, y_zip, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y, y_negbin, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y, y_zinegbin, stat = "prop_zero", binwidth = 0.005)

```

## ability to generate max 
Both negative binomial and zero-inflated negative binomial has some crazy outliers here, but looks pretty reasonable. 

```{r}

ppc_stat(y, y_zip, stat = "max")
ppc_stat(y, y_negbin, stat = "max")
ppc_stat(y, y_zinegbin, stat = "max")

```

## by group

They all predict "too much" of the same for each group (although might be good that they are more sceptical).
Best model seems to be the negative binomial. 

```{r}

ppc_stat_grouped(y, y_zip, group = d$condition_fct, stat = "prop_zero") # predicts same for groups 
ppc_stat_grouped(y, y_negbin, group = d$condition_fct, stat = "prop_zero")
ppc_stat_grouped(y, y_zinegbin, group = d$condition_fct, stat = "prop_zero")

```

# check summary
Only "interesting" thing here is that large teamsize seems to have a stronger effect for replication studies. 
Probably nothing (as we will see later) & definitely not significant. 
However, there has been a lot of attention on the large replication studies (e.g. ManyLabs). 

```{r}

summary(zip_post)
summary(negbin_post)
summary(zinegbin_post)

```

# plot model fit (focus on negative binomial)
## notes
### effect of interaction probably driven by the one outlier (more data and/or robust model?). 
### could also consider doing something else to teamsize (e.g. log) - would make sense that it is not a linear relationship.
### also consider what McElreath does with forcing it through origin. 
### add_fitted_draws() is now add_epred_draws(). 

https://mjskay.github.io/tidybayes/reference/add_predicted_draws.html

## fitted draws 
Clear that there is a lot of weight for two large studies, 
where the "control" is not cited & the experiment is cited a lot. 
Really seems to drive the (probably spurious) effect. 

```{r}

# with draws 
d %>%
  group_by(condition_fct) %>%
  data_grid(teamsize_scaled = seq_range(teamsize_scaled, n = 101)) %>%
  add_epred_draws(negbin_post, ndraws = 100) %>% # same as fitted 
  ggplot(aes(x = teamsize_scaled, y = c_5, color = ordered(condition_fct))) +
  geom_line(aes(y = .epred, group = paste(condition_fct, .draw)), alpha = 0.25) +
  geom_point(data = d)

```

## fitted CI 

```{r}

d %>%
  group_by(condition_fct) %>%
  data_grid(teamsize_scaled = seq_range(teamsize_scaled, n = 101)) %>%
  add_epred_draws(negbin_post) %>% # same as fitted 
  ggplot(aes(x = teamsize_scaled, y = c_5, color = ordered(condition_fct))) +
  stat_lineribbon(aes(y = .epred), .width = c(.95, .8, .5), alpha = 0.25) +
  geom_point(data = d) + 
  scale_fill_brewer(palette = "Greys")

```

## predicted draws
Zero is the most common value for almost any value of teamsize & condition. 
Not super informative here. 
Perhaps more interesting with draws? 

```{r}

# does a poor job predicting (but makes sense that 0 is always the value to predict...)
d %>%
  group_by(condition_fct) %>%
  data_grid(teamsize_scaled = seq_range(teamsize_scaled, n = 101)) %>%
  add_predicted_draws(negbin_post) %>% # same as fitted 
  ggplot(aes(x = teamsize_scaled, y = c_5, color = ordered(condition_fct))) +
  stat_lineribbon(aes(y = .prediction), .width = c(.8, .5), alpha = 0.25) +
  geom_point(data = d) + 
  scale_fill_brewer(palette = "Greys")

```

# HDI for parameters

```{r}

mcmc_hdi <- function(fit, title){
  
  mcmc_areas(
    fit,
    pars = c("b_condition_fctexperiment",
             "b_condition_fctcontrol",
             "b_condition_fctexperiment:teamsize_scaled",
             "b_condition_fctcontrol:teamsize_scaled"),
    prob = 0.8, # 80% intervals
    prob_outer = 0.99, # 99%
    point_est = "mean") + 
    theme_minimal() + 
    ggtitle(title)
  
}

```

```{r}

mcmc_hdi(negbin_post, "negative binomial HDI")

```

# model comparison 
consider criterion, i.e. waic, loo, bayes_R2 (I think the latter is for stacking weights?). 
best at the moment is actually the zero-inflated negative binomial, but very close to negative binomial.
should probably keep both in mind moving forward (as they both looked reasonable). 

```{r}
# add criterions 
zip_post <- add_criterion(zip_post, 
                          criterion = c("loo", "bayes_R2"))

negbin_post <- add_criterion(negbin_post, 
                             criterion = c("loo", "bayes_R2"))

zinegbin_post <- add_criterion(zinegbin_post, 
                               criterion = c("loo", "bayes_R2")) # one problematic observation.
# run loo compare
loo_compare(zip_post,
            negbin_post,
            zinegbin_post)

# also uses stacking (and gives the same as pyMC3)
loo_model_weights(zip_post,
                  negbin_post,
                  zinegbin_post)

```

# Check influential observations (pareto k) 
## check recoded (of rethinking)
