---
title: "fit interactions"
author: "Victor Møller Poulsen"
output: html_document
---

# install/read packages and source functions from fun_helper.R

```{r, include=FALSE}

# consider pacman
install.packages("pacman") 
library(pacman)

p_load(tidyverse, brms, ggthemes, bayesplot, cowplot, tidybayes, modelr)

# set up cmdstanr 
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
library(cmdstanr)
install_cmdstan(cores = 2, overwrite = TRUE)

setwd("/work/50114/MAG/modeling/code")
source("fun_helper.R")

```

# read data

```{r, include=FALSE}

# check data
d <- read_csv("/work/50114/MAG/data/modeling/psych_replication_matched.csv") %>%
  mutate(log_teamsize = log(n_authors), 
         condition_coded = ifelse(condition == "experiment", 1, 0),
         condition_fct = as_factor(condition), 
         teamsize_scaled = (n_authors-min(n_authors))/(max(n_authors)-min(n_authors)),
         days_after_2010_scaled = days_after_2010/max(days_after_2010),
         id_fct = as_factor(PaperId)) %>% # because min = 0
  glimpse()

```

# models (just focusing on negative binomial here). 
## (1) does intercept make sense when we only have one obs. in each category?
## (2) should the random effect be on the "control + experiment" group? 
## (3) random slopes out of the window, right (since we are not tracking studies over time)...

```{r}

# just simple model here (still ignoring days)
f_baseline <- bf(c_5 ~ 0 + condition_fct + condition_fct:teamsize_scaled) 
f_team_int <- bf(c_5 ~ 0 + condition_fct + condition_fct:teamsize_scaled + (1|id_fct))
f_team_full <- bf(c_5 ~ 0 + condition_fct * teamsize_scaled + (1|id_fct))
f_team_days <- bf(c_5 ~ 0 + condition_fct + condition_fct:teamsize_scaled + condition_fct:days_after_2010_scaled + (1|id_fct))

```

# get prior 
## running both negative binomial and zero-inflated

```{r, include=FALSE}

get_prior(
  formula = f_baseline,
  data = d,
  family = negbinomial()
)

get_prior(
  formula = f_baseline,
  data = d,
  family = zero_inflated_negbinomial()
)

get_prior(
  formula = f_team_int,
  data = d,
  family = negbinomial()
)

get_prior(
  formula = f_team_int,
  data = d,
  family = zero_inflated_negbinomial()
)

get_prior(
  formula = f_team_full,
  data = d,
  family = negbinomial()
)

get_prior(
  formula = f_team_full,
  data = d,
  family = zero_inflated_negbinomial()
)

get_prior(
  formula = f_team_days,
  data = d,
  family = negbinomial()
)

get_prior(
  formula = f_team_days,
  data = d,
  family = zero_inflated_negbinomial()
)

```

# set priors 

```{r}

# negbin baseline
negbin_baseline <- c(prior(gamma(0.01, 0.01), class = shape), 
                     prior(normal(0, 1), class = b))

# zinegbin baseline
zinegbin_baseline = c(prior(gamma(0.01, 0.01), class = shape), 
                      prior(normal(0, 1), class = b),
                      prior(beta(2, 2), class = zi))

# can be used for all interactions (without thinking)
negbin_interaction <- c(prior(gamma(0.01, 0.01), class = shape), 
                     prior(normal(0, 1), class = b),
                     prior(normal(0, 1), class = sd)) # a wild guess

# can be used for all interactions (without thinking)
zinegbin_interaction = c(prior(gamma(0.01, 0.01), class = shape), 
                      prior(normal(0, 1), class = b),
                      prior(beta(2, 2), class = zi),
                      prior(normal(0, 1), class = sd))



```

# fit prior models 

```{r, include=FALSE}

## baseline 
negbin_prior_baseline <- fit_model(
  family = negbinomial(), 
  formula = f_baseline,
  prior = negbin_baseline,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_baseline"
)

zinegbin_prior_baseline <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_baseline,
  prior = zinegbin_baseline,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/zinegbin_prior_baseline"
)

## teamsize interaction
negbin_prior_team_int <- fit_model(
  family = negbinomial(), 
  formula = f_team_int,
  prior = negbin_interaction,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_team_int"
)

zinegbin_prior_team_int <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_team_int,
  prior = zinegbin_interaction,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/zinegbin_prior_team_int"
)

## teamsize main + interaction
negbin_prior_team_full <- fit_model(
  family = negbinomial(), 
  formula = f_team_full,
  prior = negbin_interaction,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_team_full"
)

zinegbin_prior_team_full <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_team_full,
  prior = zinegbin_interaction,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/zinegbin_prior_team_full"
)

## both interactions 
negbin_prior_team_days <- fit_model(
  family = negbinomial(), 
  formula = f_team_days,
  prior = negbin_interaction,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_team_days"
)

zinegbin_prior_team_days <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_team_days,
  prior = zinegbin_interaction,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/zinegbin_prior_team_days"
)

```

# check priors (before sampling everything)
we do have some divergences here, but it looks ok to me. 

```{r}

prior_check <- function(model, ndraws, title, xmax){
  
  pp_check(model, 
           ndraws = ndraws) +
    labs(title = title) + 
    theme_minimal() + 
    xlim(0, xmax)
  
} 

```

```{r}

prior_check(negbin_prior_baseline, 100, "negative binomial baseline (x cutoff: 25)", 25)
prior_check(zinegbin_prior_baseline, 100, "zi-negative binomial baseline (x cutoff: 25)", 25)
prior_check(negbin_prior_team_int, 100, "negative binomial team int (x cutoff: 25)", 25)
prior_check(zinegbin_prior_team_int, 100, "zi-negative binomial team int (x cutoff: 25", 25)
prior_check(negbin_prior_team_full, 100, "negative binomial team full (x cutoff: 25)", 25)
prior_check(zinegbin_prior_team_full, 100, "zi-negative binomial team full (x cutoff: 25)", 25)
prior_check(negbin_prior_team_days, 100, "negative binomial team & days interaction (x cutoff 25)", 25)
prior_check(zinegbin_prior_team_days, 100, "zi-negative binomial team & days interaction (x cutoff 25)", 25)

```

# fit actual models 

```{r}


## baseline 
negbin_post_baseline <- fit_model(
  family = negbinomial(), 
  formula = f_baseline,
  prior = negbin_baseline,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_baseline"
)

zinegbin_post_baseline <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_baseline,
  prior = zinegbin_baseline,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/zinegbin_post_baseline"
)

## teamsize interaction
negbin_post_team_int <- fit_model(
  family = negbinomial(), 
  formula = f_team_int,
  prior = negbin_interaction,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_team_int"
)

zinegbin_post_team_int <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_team_int,
  prior = zinegbin_interaction,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/zinegbin_post_team_int"
)

## teamsize main + interaction
negbin_post_team_full <- fit_model(
  family = negbinomial(), 
  formula = f_team_full,
  prior = negbin_interaction,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_team_full"
)

zinegbin_post_team_full <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_team_full,
  prior = zinegbin_interaction,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/zinegbin_post_team_full"
)

## both interactions 
negbin_post_team_days <- fit_model(
  family = negbinomial(), 
  formula = f_team_days,
  prior = negbin_interaction,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_team_days"
)

zinegbin_post_team_days <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f_team_days,
  prior = zinegbin_interaction,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/zinegbin_post_team_days"
)


```

