---
title: "Test Interaction"
author: "Victor Møller Poulsen"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
# consider pacman
install.packages("pacman") 
library(pacman)

p_load(tidyverse, brms, ggthemes, bayesplot, cowplot, tidybayes, modelr)

# set up cmdstanr 
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
library(cmdstanr)
install_cmdstan(cores = 2, overwrite = TRUE)

setwd("/work/50114/MAG/modeling/code")
source("fun_helper.R")
```

```{r}

# check data
d <- read_csv("/work/50114/MAG/data/modeling/psych_replication_matched.csv") %>%
  mutate(log_teamsize = log(n_authors), 
         condition_coded = ifelse(condition == "experiment", 1, 0),
         condition_fct = as_factor(condition), 
         teamsize_scaled = (n_authors-min(n_authors))/(max(n_authors)-min(n_authors)),
         days_after_2010_scaled = days_after_2010/max(days_after_2010),
         id_fct = as_factor(PaperId)) %>% # because min = 0
  glimpse()

```

Different ways of specifying something similar. 
We had issues with model (f_team_0) earlier. 
Trying to troubleshoot whether it is related to intercept & 
I think that (0 + Intercept) syntax is actually more appropriate 
since it does not assume mean centering (something like that). 

```{r}

f_team_0 <- bf(c_5 ~ 0 + condition_fct + condition_fct:teamsize_scaled + (1|id_fct))
f_team_1 <- bf(c_5 ~ 1 + condition_fct + condition_fct:teamsize_scaled + (1|id_fct))
f_team_01 <- bf(c_5 ~ 0 + Intercept + condition_fct + condition_fct:teamsize_scaled + (1|id_fct))

```

Just doing negbinomial() for now, since we had Rhat issues for both negative binomial and
zero-inflated negative binomial (does not seem to be the main cause of issues). 

f_team_0: b, sd, shape
f_team_1: b, Intercept, sd, shape
f_team_01: b, sd, shape (Intercept becomes b). 

```{r, include=FALSE}

get_prior(
  formula = f_team_0,
  data = d,
  family = negbinomial()
)

get_prior(
  formula = f_team_1,
  data = d,
  family = negbinomial()
)

get_prior(
  formula = f_team_01,
  data = d,
  family = negbinomial()
)

```

set priors

```{r}

# negbin baseline
negbin_0 <- c(prior(gamma(0.01, 0.01), class = shape), 
              prior(normal(0, 1), class = b),
              prior(normal(0, 1), class = sd)) # a wild guess

# zinegbin baseline
negbin_1 = c(prior(gamma(0.01, 0.01), class = shape), 
             prior(normal(0, 1), class = b),
             prior(normal(0, 1), class = Intercept),
             prior(normal(0, 1), class = sd)) # a wild guess

# can be used for all interactions (without thinking)
negbin_01 <- c(prior(gamma(0.01, 0.01), class = shape), 
              prior(normal(0, 1), class = b),
              prior(normal(0, 1), class = sd)) # a wild guess



```

# sample prior only 
Some warnings and divergences. 

```{r, include=FALSE}

## baseline 
negbin_prior_0 <- fit_model(
  family = negbinomial(), 
  formula = f_team_0,
  prior = negbin_0,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_0"
)

## baseline 
negbin_prior_1 <- fit_model(
  family = negbinomial(), 
  formula = f_team_1,
  prior = negbin_1,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_1"
)

## baseline 
negbin_prior_01 <- fit_model(
  family = negbinomial(), 
  formula = f_team_01,
  prior = negbin_01,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_01"
)



```

# check priors

```{r}

prior_check <- function(model, ndraws, title, xmax){
  
  pp_check(model, 
           ndraws = ndraws) +
    labs(title = title) + 
    theme_minimal() + 
    xlim(0, xmax)
  
} 

```

```{r}

prior_check(negbin_prior_0, 100, "No Intercept (x cutoff: 25)", 25)
prior_check(negbin_prior_1, 100, "Intercept (x cutoff: 25)", 25)
prior_check(negbin_prior_01, 100, "0 + Intercept (x cutoff: 25)", 25)

```

# fit models 

```{r}

## baseline 
negbin_post_0 <- fit_model(
  family = negbinomial(), 
  formula = f_team_0,
  prior = negbin_0,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_0"
)

## baseline 
negbin_post_1 <- fit_model(
  family = negbinomial(), 
  formula = f_team_1,
  prior = negbin_1,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_1"
)

## baseline 
negbin_post_01 <- fit_model(
  family = negbinomial(), 
  formula = f_team_01,
  prior = negbin_01,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_01"
)


```

# check traces

```{r}

# some auto-correlation
# effect almost entirely in random effect
plot(negbin_post_0, N = 3)

```


```{r}

# some auto-correlation
# effect almost entirely in random effect
plot(negbin_post_1, N = 3)

```

```{r}

# same issues 
plot(negbin_post_01, N = 3)

```

Seems like the issue is not the particular specification: 
This post explains why it is problematic to have random effect for single observation: 
https://stats.stackexchange.com/questions/242821/how-will-random-effects-with-only-1-observation-affect-a-generalized-linear-mixe

# Try with random effect per group (i.e. experiment + control) 

## create the variable (as factor)

```{r}

d <- d %>%
  mutate(id_match = as_factor(match_group))

```

# new model formulae

```{r}

f_team_match_0 <- bf(c_5 ~ 0 + condition_fct + condition_fct:teamsize_scaled + (1|id_match))
f_team_match_1 <- bf(c_5 ~ 1 + condition_fct + condition_fct:teamsize_scaled + (1|id_match))
f_team_match_01 <- bf(c_5 ~ 0 + Intercept + condition_fct + condition_fct:teamsize_scaled + (1|id_match))

```

# prior 

```{r}

## baseline 
negbin_prior_match_0 <- fit_model(
  family = negbinomial(), 
  formula = f_team_match_0,
  prior = negbin_0, # same prior 
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_match_0"
)

## baseline 
negbin_prior_match_1 <- fit_model(
  family = negbinomial(), 
  formula = f_team_match_1,
  prior = negbin_1,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_match_1"
)

## baseline 
negbin_prior_match_01 <- fit_model(
  family = negbinomial(), 
  formula = f_team_match_01,
  prior = negbin_01,
  sample_prior = "only",
  file = "/work/50114/MAG/modeling/models/negbin_prior_match_01"
)

```

# prior check 
A few divergent transitions (for all of them, between 7-9/4000)

```{r}

prior_check(negbin_prior_0, 100, "No Intercept (x cutoff: 25)", 25)
prior_check(negbin_prior_1, 100, "Intercept (x cutoff: 25)", 25)
prior_check(negbin_prior_01, 100, "0 + Intercept (x cutoff: 25)", 25)

```

# fit models 
a few pareto_k > 0.7 (but few compared to above). 

```{r}
## baseline 
negbin_post_match_0 <- fit_model(
  family = negbinomial(), 
  formula = f_team_match_0,
  prior = negbin_0,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_match_0"
)

## baseline 
negbin_post_match_1 <- fit_model(
  family = negbinomial(), 
  formula = f_team_match_1,
  prior = negbin_1,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_match_1"
)

## baseline 
negbin_post_match_01 <- fit_model(
  family = negbinomial(), 
  formula = f_team_match_01,
  prior = negbin_01,
  sample_prior = TRUE,
  file = "/work/50114/MAG/modeling/models/negbin_post_match_01"
)

```

# traces
Looks MUCH better now than before. 

```{r}

# looks pretty good. 
plot(negbin_post_match_0, N = 3)

```

```{r}

# same as above 
plot(negbin_post_match_1, N = 3)
```

```{r}

# slightly different: might be more appropriate? 
plot(negbin_post_match_01, N = 3)

```

# check estimates
More effective samples for intercept models.
No Rhat issues, looks pretty good. 
Not strictly "significant" given 95% CI, but close -- 
and also stronger effect of teamsize (although problematic because of outlier) --
connected with the pareto k issue. 

```{r}

print(negbin_post_match_0)
print(negbin_post_match_1) # more effective samples 
print(negbin_post_match_01) 

```

# plot implications

```{r}

y <- d$c_5
y_0 <- posterior_predict(negbin_post_match_0, draws = 500)
y_1 <- posterior_predict(negbin_post_match_1, draws = 500)
y_01 <- posterior_predict(negbin_post_match_01, draws = 500)

```

## no intercept 
looks pretty good. 
a lot of uncertainty around 0 and 1 still. 

```{r}
ppc_dens_overlay(y, y_0[1:50, ]) + xlim(0, 50)
ppc_hist(y, y_0[1:5, ]) + xlim(-1, 50)
```

## regular intercept
Perhaps a bit worse (e.g. with the undershoot at around c_5 = 5). 

```{r}
ppc_dens_overlay(y, y_1[1:50, ]) + xlim(0, 50)
ppc_hist(y, y_1[1:5, ]) + xlim(-1, 50)
```

## 0 + Intercept
Looks more or less the same.

```{r}
ppc_dens_overlay(y, y_01[1:50, ]) + xlim(0, 50)
ppc_hist(y, y_01[1:5, ]) + xlim(-1, 50)
```

# pareto k issues 
https://bookdown.org/content/4857/monsters-and-mixtures.html

## no intercept 
mainly studies that are (relatively) low teamsize and high citation
```{r}

d %>% 
  mutate(k = negbin_post_match_0$criteria$loo$diagnostics$pareto_k) %>% 
  filter(k > .7) %>% 
  select(c_5, teamsize_scaled, condition_fct, id_match, k)

```

## 1 + ...
some of the same here, mainly studies with high c_5 and low teamsize. 

```{r}

d %>% 
  mutate(k = negbin_post_match_1$criteria$loo$diagnostics$pareto_k) %>% 
  filter(k > .7) %>% 
  select(c_5, teamsize_scaled, condition_fct, id_match, k)

```


## 0 + Intercept 
Seems to handle influential observations much better.
Good explanation for all of the outliers: 
(1) the two control studies are high influence because they are the same (fix earlier in pipeline)
(2) the experiment study is high influence because there are very few studies with high teamsize (could do log-transformation) and the corresponding control study has zero citations... so this has an outsize influence on the interaction with condition and teamsize. 

```{r}

# two studies that are the same in control (issue to be resolved earlier in the pipeline).
# the outlier study (experiment) which is max in teamsize and also extremely high citation
# whereas the 
d %>% 
  mutate(k = negbin_post_match_01$criteria$loo$diagnostics$pareto_k) %>% 
  filter(k > .7) %>% 
  select(c_5, teamsize_scaled, condition_fct, id_match, k)

```

# Quick model comparison
Basically no difference, but appears to prefer the intercept models. 
Do we know why that is? 

```{r}

loo_compare(negbin_post_match_0,
            negbin_post_match_1,
            negbin_post_match_01)

loo_model_weights(negbin_post_match_0,
                  negbin_post_match_1,
                  negbin_post_match_01)
```



