---
title: "EDA"
author: "Victor M. Poulsen" 
output: html_document
---

# Set-up

```{r setup, include=FALSE}

# consider pacman
if (!require("pacman")){
  install.packages("pacman") # repos = "http://cran.r-project.org"
}

library(pacman)
p_load(tidyverse, 
       cowplot,
       latex2exp,
       ggdist,
       ggpubr,
       gridExtra)

```


# Aesthetic set-up

```{r}

color_map <- c("experiment" = '#e6550d',
               "control" = '#fdae6b')

color_lst <- c('#e6550d', '#fdae6b')

theme_set(theme_classic())
title = 18
label = 14
tick = 12 # same as legend

```

# PATHS

```{r}

inpath <- "/work/50114/MAG/data/modeling/"
infile <- "psychology_replicat_matched.csv" # "psychology_replication_matched.csv"
filename <- "replication_query" #"replication_fos"
outpath_csv <- "/work/50114/MAG/data/EDA/"
outpath_fig <- "/work/50114/MAG/fig/EDA/"

```


# Read data

```{r}

d <- read_csv(paste0(inpath, infile)) %>% 
  mutate(log_teamsize = log(n_authors), 
         condition_fct = as_factor(condition), 
         condition_fct = fct_relevel(condition_fct, c("experiment", "control")),
         id_match = as_factor(match_group),
         id_dct = as_factor(PaperId),
         year_after_2005 = Year - 2005) %>%
  select(condition_fct, condition, n_authors, log_teamsize, Year, year_after_2005, c_5)

```

# c5

## prep 

```{r}

max_count <- d %>% group_by(condition_fct, c_5) %>%
  summarize(count = n()) %>% 
  ungroup() %>%
  summarize(max = max(count))

```

```{r}

x_max <- ceiling(max(max_count[[1]])/10)*10

```


## histogram 

```{r}

p_c5_hist <- d %>% ggplot(aes(x = c_5, fill = condition_fct)) + 
    geom_bar(stat = "count", width = 0.8) + 
    facet_wrap(~condition_fct) +
    lims(x = c(-0.5, 50), y = c(0, x_max)) +
  theme(axis.text = element_text(size = tick), #element_blank(), # element_text(size = tick),
        axis.title = element_text(size = label),
        legend.text = element_text(size = tick),
        legend.position = "bottom",
        strip.background = element_blank(),
        strip.text = element_blank()) + # element_text(size = label)) +
  scale_fill_manual(values = color_map,
                    guide = guide_legend(title = NULL)) + 
  labs(x = TeX("$c_5$"),
       y = "count")
```


## halfeye + box

```{r}

p_c5_halfeye <- d %>% ggplot(aes(c_5, condition_fct, fill = condition_fct)) + 
  ggdist::stat_halfeye(point_color = NA, .width = 0, height = 1) +  
  geom_boxplot(width = .1, outlier.shape = NA, position = position_nudge(y = -0.1), show.legend=FALSE) + 
  scale_fill_manual(values = color_map,
                    guide = guide_legend(title = NULL)) + 
  scale_color_manual(values = color_map,
                     guide = guide_legend(title = NULL)) + 
  #xlim(0, 300) +
  scale_y_discrete(expand = c(0.12, 0.12)) +
  scale_x_continuous(limits = c(0, 300), expand = c(0.02, 0.02)) + 
  labs(#title = TeX("Outcome Distribution ($c_5$)"),
       y = "condition", 
       x = TeX("$c_5$")) + 
  theme(#plot.title = element_text(hjust = 0.5, size = title),
        axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        legend.text = element_text(size = tick),
        legend.position = "none",
        #axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

```

## gather 

```{r}

p_grid <- ggarrange(p_c5_halfeye, p_c5_hist, ncol = 1,
          labels = c("A", "B")) 

```

## annotate

```{r}

#p_annotate <- annotate_figure(p_grid, top = text_grob(TeX("Outcome Distribution ($c_5$)"), size #= title))

```

## save 

```{r}

ggsave(filename = paste0(outpath_fig, filename, "_c5_distribution.pdf"), 
       plot = p_grid,
       width = 8,
       height = 6)

```

# teamsize 
## NB: we put less weight on the really far end of the tail (e.g. those with more than 100 citations) & 
## more emphasize the difference between small and medium sized studies. 
## this is with 95% confidence interval around the mean linear trend. 
## setting ylim(0, 300) which excludes one highly cited study for the benefit of actually being able to see anything... 

## annotated plot 

```{r}

x_low = 1
x_mid = 2
x_high = 4
x_top = 8
x_space = 0.05
y_main = 280
y_space = 3
size = 0.2

p_teamsize_log <- d %>% ggplot(aes(x = log_teamsize, y = c_5, color = condition_fct)) + 
  geom_jitter(height = 0, size = 1.5, alpha = 0.8) + 
  geom_smooth(method = "lm", aes(fill = condition_fct)) + 
  ylim(0, 300) + 
  scale_colour_manual(values = color_map,
                      guide = guide_legend(title = NULL)) +
  scale_fill_manual(values = color_map,
                    guide = guide_legend(title = NULL)) + 
  geom_segment(aes(x = log(x_low), # low horizontal
                   y = y_main, 
                   xend = log(x_mid - x_space), 
                   yend = y_main), 
               color = 'black', size = size) +
  geom_segment(aes(x = log(x_high + x_space), # high horizontal
                   y = y_main, 
                   xend = log(x_top), 
                   yend = y_main), 
               color = 'black', size = size) + 
  geom_segment(aes(x = log(x_low), # low vertical (left)
                   y = y_main - y_space, 
                   xend = log(x_low), 
                   yend = y_main + y_space), 
               color = 'black', size = size) +
  geom_segment(aes(x = log(x_mid - x_space), # low vertical (right)
                   y = y_main - y_space, 
                   xend = log(x_mid - x_space), 
                   yend = y_main + y_space), 
               color = 'black', size = size) + 
  geom_segment(aes(x = log(x_high + x_space), # high vertical (left)
                   y = y_main - y_space, 
                   xend = log(x_high + x_space), 
                   yend = y_main + y_space), 
               color = 'black', size = size) +
  geom_segment(aes(x = log(x_top),
                   y = y_main - y_space,
                   xend = log(x_top),
                   yend = y_main + y_space),
               color = 'black', size = size) +
  annotate("text", x = log(1.4), y = y_main + 20, label = TeX("$TEAMSIZE \\in (1, 2)$"), size = 4) + 
  annotate("text", x = log(5.9), y = y_main + 20, label = TeX("$TEAMSIZE \\in (4, 8)$"), size = 4) + 
  labs(title = "",
       x = "log(TEAMSIZE)",
       y = TeX("$c_5$")) +
  theme(axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        legend.text = element_text(size = tick),
        legend.position = "bottom") 

p_teamsize_log

```

## base plot

```{r}

p_teamsize_raw <- d %>% ggplot(aes(x = n_authors, y = c_5, color = condition_fct)) + 
  geom_jitter(height = 0, size = 1.5, alpha = 0.8) + 
  geom_smooth(method = "lm", 
              aes(fill = condition_fct)) + 
  ylim(0, 300) + 
  scale_colour_manual(values = color_map,
                      guide = guide_legend(title = NULL)) +
  scale_fill_manual(values = color_map,
                    guide = guide_legend(title = NULL)) + 
  labs(title = "",
       x = "TEAMSIZE",
       y = TeX("$c_5$")) + 
  theme(axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        #legend.text = element_text(size = tick),
        legend.position = "none")

```


# year 
## again excludes this one study to be able to actually see what is going on. 

```{r}

p_year <- d %>% ggplot(aes(x = Year, y = c_5, color = condition_fct)) + 
  geom_jitter(height = 0, width = 0.2, size = 1.5, alpha = 0.8) + 
  geom_smooth(method = "lm",
              aes(fill = condition_fct)) + 
  ylim(0, 300) + 
  scale_colour_manual(values = color_map,
                      guide = guide_legend(title = NULL)) +
  scale_fill_manual(values = color_map,
                    guide = guide_legend(title = NULL)) + 
  labs(title = "",
       x = "YEAR",
       y = TeX("$c_5$")) + 
  theme(axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        #legend.text = element_text(size = tick),
        legend.position = "none")

```

## gather in grid 

```{r}

gt <- arrangeGrob(p_teamsize_log, 
                  p_year,
                  p_teamsize_raw, 
                  ncol = 2, 
                  nrow = 2, 
                  layout_matrix = rbind(c(2,3), c(1,1),
                                        heights=c(4,1)))

# Add labels to the arranged plots
p_grid <- as_ggplot(gt) +                                # transform to a ggplot
  draw_plot_label(label = c("A", "B", "C"), size = 15,
                  x = c(0, 0.5, 0), y = c(1, 1, 0.65)) # Add labels


```

## annotate

```{r}

#p_annotate <- annotate_figure(p_grid, top = text_grob(TeX("Predictors (EDA)"), size = title))

```

## ggsave

```{r}

ggsave(filename = paste0(outpath_fig, filename, "_c5_predictors.pdf"), 
       plot = p_grid,
       width = 8,
       height = 6)

```


## Summary statistics

## NOTE on LOG

```{r}

# same distance
log(2) - log(1) # 0.693
log(4) - log(2) # 0.693
log(8) - log(4) # 0.693
log(32) - log(16) # 0.693
log(128) - log(64) # 0.693

```

## Mean, Median, Max, standard deviation (sd), Number of Zero's

```{r}

# divisor 
n_each_category = nrow(d)/2

# summary statistics
d_stats <- d %>% group_by(condition_fct) %>% 
  summarize(c5_mean = mean(c_5), # much greater mean
            c5_median = median(c_5),
            c5_max = max(c_5),
            c5_sd = sd(c_5)) # also much greater sd 
  
# number of zero's
d_zero <- d %>% filter(c_5 == 0) %>% 
  group_by(condition_fct) %>% 
  summarize(c5_zero = n()) %>% 
  mutate(c5_zero_fraction = c5_zero / n_each_category)

# join it together 
d_summary <- merge(d_stats, d_zero, by = "condition_fct")

# write file 
write_csv(d_summary, paste0(outpath, filename, "d_summary.csv"))

```

#### test it out for QUERY ####

```{r}

d_zero

```

```{r}
d_stats
```



