---
title: "Untitled"
output: html_document
---

```{r, include=FALSE}

# consider pacman
if (!require("pacman")){
  install.packages("pacman") # repos = "http://cran.r-project.org"
}

library(pacman)
p_load(tidyverse, brms, ggthemes, bayesplot, cowplot, tidybayes, modelr, latex2exp, ggpubr)

# set up cmdstanr if it is not already present
if (!require('cmdstanr')){
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
library(cmdstanr)
install_cmdstan(cores = 2, overwrite = TRUE)
}


```

# FOS 

```{r, include=FALSE}

inpath <- "/work/50114/MAG/data/modeling/"
infile <- "psychology_replication_matched.csv"

csv_path <- paste0(inpath, infile)

# check data
d_fos <- read_csv(csv_path) %>%
  mutate(log_teamsize = log(n_authors), 
         condition_fct = as_factor(condition), 
         condition_fct = fct_relevel(condition_fct, c("experiment", "control")),
         id_match = as_factor(match_group),
         id_dct = as_factor(PaperId),
         year_after_2005 = Year - 2005) 

```


```{r}

inpath_post <- "/work/50114/MAG/modeling/models/replication_fos/main_model/m_post.rds"
m_fos <- readRDS(paste0(inpath_post))

```


# EVIDENCE RATIO

## mean values of YEAR and TEAMSIZE

```{r}

d_sum <- d_fos %>% summarize(mean_ts = mean(log_teamsize), # 0.9076
                mean_ts_nat = exp(mean_ts), # 2.478
                mean_y = mean(year_after_2005), # 6.244
                mean_y_nat = mean_y + 2005) %>% # 2011.24
  glimpse()

# save
write_csv(d_sum, paste0(outpath, "data_statistics.csv"))

```

```{r}

d_fos

```


Our group-difference for the INTERCEPT parameter is the expected difference in citations between
replication (experiment) and non-replication (control) studies for a solo-authored paper published
in 2005. A better idea about which group of studies (replication, control) is more cited on average
is obtained when comparing the expected citation difference for mean values of our predictors, 
TEAMSIZE and YEAR. The mean TEAMSIZE in the R_{FOS} data set is 2.48 authors and the mean YEAR 
in our data is 2011.24. When comparing the inferred parameters for replication and control studies
with these values, the expected citation difference is significant (Est: $3.3$, CIs: $1.6, 5.07$, Evid.Ratio $> 1000$, Post.Prob = $1$). All posterior probability is consistent with replication (experiment) studies being cited more
than non-replication (control) studies for mean values of TEAMSIZE and YEAR.

```{r}

d_means <- hypothesis(m_fos, glue("exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.908 + condition_fctexperiment:year_after_2005*{val}) - exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.908 + condition_fctcontrol:year_after_2005*6.24) > 0"), 
alpha = 0.05)
d_means # Est: 3.3 (1.6, 5.07), Evid.Ratio = 7999, Post.Prob = 1

```

```{r}

q_cont <- hypothesis(m_fos, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.908 + condition_fctcontrol:year_after_2005*6.24) > 0", alpha = 0.05)
q_cont # Est: 8.04 (6.89, 9.28)

```

```{r}

q_cont <- hypothesis(m_fos, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.908 + condition_fctexperiment:year_after_2005*6.24) > 0", alpha = 0.05)
q_cont # Est: 11.35 (9.79, 13.01)

```

We visualize data generated

```{r}

d_fos_mean <- d_fos %>%
    group_by(condition_fct) %>%
    data_grid(log_teamsize = mean(log_teamsize), # sequence of values
              year_after_2005 = mean(year_after_2005)) %>% # fixed year
    add_epred_draws(m_post, re_formula = NA) # no random eff. uncertainty

```

```{r}

color_map <- c("experiment" = '#e6550d',
               "control" = '#fdae6b')
label = 14
tick = 12

d_fos_mean %>% ggplot(aes(.epred, condition_fct, fill = condition_fct)) + 
  ggdist::stat_halfeye(point_color = NA, .width = 0, height = 1) +  
  geom_boxplot(width = .1, outlier.shape = NA, position = position_nudge(y = -0.1), show.legend=FALSE) + 
  scale_fill_manual(values = color_map,
                    guide = guide_legend(title = NULL)) + 
  scale_color_manual(values = color_map,
                     guide = guide_legend(title = NULL)) + 
  #xlim(0, 300) +
  scale_y_discrete(expand = c(0.12, 0.12)) +
  scale_x_continuous(limits = c(5, 15), expand = c(0.02, 0.02)) + 
  labs(#title = TeX("Outcome Distribution ($c_5$)"),
       y = "condition", 
       x = TeX("$c_5$")) + 
  theme(#plot.title = element_text(hjust = 0.5, size = title),
        axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        legend.text = element_text(size = tick),
        legend.position = "none",
        #axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

```

## predictors (YEAR)

```{r}

d_year <- hypothesis(m_fos, "(exp(condition_fctexperiment + condition_fctexperiment:year_after_2005*10) - 
                         exp(condition_fctexperiment)) - (exp(condition_fctcontrol + condition_fctcontrol:year_after_2005*10) - exp(condition_fctcontrol)) > 0", alpha = 0.05)

d_year # 8.42 (4.81, 12.5), Evid.Ratio: INF, Post.Prob = 1


```

```{r}

d_year_exp <- hypothesis(m_fos, "exp(condition_fctexperiment + condition_fctexperiment:year_after_2005*10) - 
                         exp(condition_fctexperiment) > 0", alpha = 0.05)
d_year_exp # Est: 5.57 (2.46, 9.1), Evid.Ratio = 614, Post.Prob = 1.

```

```{r}

d_cont_exp <- hypothesis(m_fos, "exp(condition_fctcontrol + condition_fctcontrol:year_after_2005*10) - 
                         exp(condition_fctcontrol) > 0", alpha = 0.05)
d_cont_exp # Est: -2.85 (-4.76, -1.19), Evid.Ratio: 0, Post.Prob = 0. 

```

```{r}

conditions <- data.frame(log_teamsize = 0)
plot(conditional_effects(m_fos, effects = "year_after_2005:condition_fct", conditions = conditions))

```

## plot 

```{r}

```


## predictors (TEAMSIZE)

```{r}
log(2) # 0.693
log(5) # 1.609
log(20) # 2.996
```


```{r}

d_team_one_e <- hypothesis(m_fos, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) - 
                         exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.693) > 0", alpha = 0.05)
d_team_one_e # Est: 2.23 (0.92, 3.84), Evid = 1599, Post.Prob = 1

```

```{r}

d_team_two_e <- hypothesis(m_fos, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*2.996) - 
                         exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) > 0", alpha = 0.05)
d_team_two_e # Est: 5.03 (1.65, 9.85), Evid = 1599, Post.Prob = 1

```

```{r}

d_team_one_c <- hypothesis(m_fos, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.693) > 0", alpha = 0.05)
d_team_one_c # Est: 11.22 (7.43, 15.83), Evid.Ratio=Inf, Post.Prob=1

```

```{r}

d_team_two_c <- hypothesis(m_fos, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*2.996) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) > 0", alpha = 0.05)
d_team_two_c # Est: 45.59 (24.4, 75.24), Evid.Ratio=Inf, Post.Prob=1


```

```{r}
d_team_cont_1 <- hypothesis(m_fos, "(exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.693)) - (exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) - exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.693)) > 0", alpha = 0.05)
d_team_cont_1 # Est: 8.99 (4.89, 13.69), Evid: INF, post.prob = 1

```


```{r}
d_team_cont_2 <- hypothesis(m_fos, "(exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*2.996) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609)) - (exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*2.996) - exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609)) > 0", alpha = 0.05)
d_team_cont_2 # Est: 40.56 (18.78, 70.23), Evid.Ratio = INF, Post.Prob = 1 
```


# Q query (robustness)

We compare the expected citation difference for mean values of our predictors (TEAMSIZE, YEAR) for 
R_{QUERY} as a robustness check. Recall that for the $R_{FOS}$ model the expected number of citations
was significantly higher for replication studies as compared to matched controls. This result is 
replicated for the model conditioned on $R_{QUERY}$ data, and the effect is found the difference is
found to be even larger (Est: $4.58$, CIs: $3.25, 5.93$, Evid.Ratio: $> 1000$, Post.Prob = $1$). In this 
model, replication studies are expected to receive $4.58$ citations more than non-replication studies for
mean values of our predictors (YEAR, TEAMSIZE). The mean values in this case are slightly different, as they
are based on a different data set, and mean TEAMSIZE is slightly higher (3.15) while YEAR is slightly lower 
(2010.77). Overall the expected number of citations is larger for each group in the data as compared to 
in the $R_{FOS}$ model. For the control group the expected $c_5$ is $9.66 (8.69, 10.68)$ while for the experiment group (replications) the expected $c_5$ is $14.24 (13.05, 15.48)$. The heterogeneity
between the two data sets and the consistency of the overall effect that replication studies are cited more
than non-replication studies suggests that this is a robust finding. 

```{r}

inpath_post <- "/work/50114/MAG/modeling/models/replication_keyword/main_model/m_post.rds"
q_post <- readRDS(paste0(inpath_post))

```

```{r}

inpath <- "/work/50114/MAG/data/modeling/"
infile <- "psychology_replicat_matched.csv"

csv_path <- paste0(inpath, infile)

# check data
q <- read_csv(csv_path) %>%
  mutate(log_teamsize = log(n_authors), 
         condition_fct = as_factor(condition), 
         condition_fct = fct_relevel(condition_fct, c("experiment", "control")),
         id_match = as_factor(match_group),
         id_dct = as_factor(PaperId),
         year_after_2005 = Year - 2005) 

```


```{r}

q %>% summarize(mean_ts = mean(log_teamsize), # 1.148 
                mean_ts_nat = exp(mean_ts), # 3.15 
                mean_y = mean(year_after_2005), # 5.77 
                mean_y_nat = mean_y + 2005) %>% # 2010.77 
  glimpse()

```

```{r}

q_means <- hypothesis(q_post, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.148 + condition_fctexperiment:year_after_2005*5.77) - exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.148 + condition_fctcontrol:year_after_2005*5.77) > 0", 
alpha = 0.05)
q_means # Est: 4.58 (3.25, 5.93) Evid = inf, Post.Prob = 1. 

```

```{r}

q_cont <- hypothesis(q_post, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.148 + condition_fctcontrol:year_after_2005*5.77) > 0", alpha = 0.05)
q_cont # Est: 9.66 (8.69, 10.68)

```

```{r}

q_exp <- hypothesis(q_post, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.148 + condition_fctexperiment:year_after_2005*5.77) > 0")
q_exp # 14.24 (13.05, 15.48)

```

```{r}

q_overall <- q %>%
    group_by(condition_fct) %>%
    data_grid(log_teamsize = mean(log_teamsize), # sequence of values
              year_after_2005 = mean(year_after_2005)) %>% # fixed year
    add_epred_draws(m_post, re_formula = NA) # no random eff. uncertainty

```

```{r}
q_overall %>% ggplot(aes(.epred, condition_fct, fill = condition_fct)) + 
  ggdist::stat_halfeye(point_color = NA, .width = 0, height = 1) +  
  geom_boxplot(width = .1, outlier.shape = NA, position = position_nudge(y = -0.1), show.legend=FALSE) + 
  scale_fill_manual(values = color_map,
                    guide = guide_legend(title = NULL)) + 
  scale_color_manual(values = color_map,
                     guide = guide_legend(title = NULL)) + 
  #xlim(0, 300) +
  scale_y_discrete(expand = c(0.12, 0.12)) +
  scale_x_continuous(limits = c(5, 20), expand = c(0.02, 0.02)) + 
  labs(#title = TeX("Outcome Distribution ($c_5$)"),
       y = "condition", 
       x = TeX("$c_5$")) + 
  theme(#plot.title = element_text(hjust = 0.5, size = title),
        axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        legend.text = element_text(size = tick),
        legend.position = "none",
        #axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

```

## predictors (YEAR)

```{r}

d_year <- hypothesis(q_post, "(exp(condition_fctexperiment + condition_fctexperiment:year_after_2005*10) - 
                         exp(condition_fctexperiment)) - (exp(condition_fctcontrol + condition_fctcontrol:year_after_2005*10) - exp(condition_fctcontrol)) > 0", alpha = 0.05)

d_year # 0 (-1.96, 1.9), Post.Prob = 0.5


```

```{r}

d_year_exp <- hypothesis(q_post, "exp(condition_fctexperiment + condition_fctexperiment:year_after_2005*10) - 
                         exp(condition_fctexperiment) < 0", alpha = 0.05)
d_year_exp # Est: -1.23 (-3.02, 0.3), Post.Prob = 0.91

```

```{r}

d_cont_exp <- hypothesis(q_post, "exp(condition_fctcontrol + condition_fctcontrol:year_after_2005*10) - 
                         exp(condition_fctcontrol) < 0", alpha = 0.05)
d_cont_exp # Est: -1.32 (-2.39, -0.32), Post.Prob = 0.99

```

## predictors (TEAMSIZE)

```{r}
log(2) # 0.693
log(5) # 1.609
log(20) # 2.996
```


```{r}

d_team_one_e <- hypothesis(q_post, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) - 
                         exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.693) > 0", alpha = 0.05)
d_team_one_e # Est: 9.35 (7.34, 11.57), Evid = INF, Post.Prob = 1

```

```{r}

d_team_two_e <- hypothesis(q_post, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*2.996) - 
                         exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) > 0", alpha = 0.05)
d_team_two_e # Est: 30.14 (21.07, 40.79), Evid = INF, Post.Prob = 1

```

```{r}

d_team_one_c <- hypothesis(q_post, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.693) > 0", alpha = 0.05)
d_team_one_c # Est: 8.15 (6.47, 10.04), Evid.Ratio=Inf, Post.Prob=1

```

```{r}

d_team_two_c <- hypothesis(q_post, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*2.996) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) > 0", alpha = 0.05)
d_team_two_c # Est: 30.43 (21.15, 41.74), Evid.Ratio=Inf, Post.Prob=1


```

```{r}
d_team_cont_1 <- hypothesis(q_post, "(exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) - 
                         exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.693)) - (exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) - exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.693)) > 0", alpha = 0.05)
d_team_cont_1 # Est: 8.99 (4.89, 13.69), Evid: INF, post.prob = 1 (higher REPLICATION)

```


```{r}
d_team_cont_2 <- hypothesis(q_post, "(exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*2.996) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609)) - (exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*2.996) - exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609)) > 0", alpha = 0.05)
d_team_cont_2 # Est: 40.56 (18.78, 70.23), Evid.Ratio = INF, Post.Prob = 1 (HIGHER CONTROL)
```
