---
title: "Explore Priors"
author: "Victor M. Poulsen" 
output: html_document
---

# Path setup

```{r setup, include=FALSE}

outpath <- "/work/50114/MAG/fig/modeling/replication_fos/explore_priors/"

```

```{r, include=FALSE}

# consider pacman
if (!require("pacman")){
  install.packages("pacman") # repos = "http://cran.r-project.org"
}

library(pacman)
p_load(tidyverse, 
       brms, 
       ggthemes, 
       bayesplot, 
       cowplot, 
       tidybayes, 
       modelr, 
       latex2exp, 
       ggpubr,
       extraDistr)

```

# Visual setup

```{r}

color_map <- c("Prior" = '#fee6ce', 
               "Posterior (control)" = '#fdae6b', 
               "Posterior (experiment)" = '#e6550d')

color_lst <- c('#fee6ce', '#fdae6b', '#e6550d')

theme_set(theme_classic())
title = 18
label = 14
tick = 12 # same as legend

```

# Shape/Phi parameter

## plot gamma: 
* For gamma() it is shape and rate (not scale I think). 
--> first argument is shape. 
--> second argument is rate: https://bookdown.org/content/4857/monsters-and-mixtures.html


```{r}

plot_gamma <- function(title_size, 
                       tick_size, 
                       label_size, 
                       shape = .01,
                       rate = .01,
                       n = 10000, 
                       from = 0, 
                       to = 10, 
                       by = .01){
  
  d <- tibble(x = rgamma(n, shape, rate)) %>% # rethinking::rgamma2(n, shape, scale)
    ggplot(aes(x = x)) +
    geom_ribbon(aes(ymin = 0,
                    ymax = dgamma(x, shape, rate)),
              fill = color_lst[2],
              alpha = 1) +
    stat_pointinterval(aes(y = 0), .width = c(.5, .95)) +
    labs(title = TeX("$\\gamma(.01, .01)$"), # Prior predictive distribution 
         x = TeX("Shape ($\\phi$)"),
         y = "") + 
    coord_cartesian(xlim = c(0, 10), ylim = (c(0, 1))) +
    theme(plot.title = element_text(size = title_size, hjust = 0.5),
          legend.title = element_blank(),
          axis.text = element_text(size = tick_size),
          axis.title = element_text(size = label_size),
          legend.text = element_text(size = tick_size)) 
  
  return(d)
}


```

## plot gamma 

```{r}

p_gamma <- plot_gamma(title_size = title,
                      tick_size = tick,
                      label_size = label)

```

## plot exp:
* For exponential(rate) it is just rate. 

```{r}

plot_exp <- function(rate, 
                     title_size, 
                     tick_size, 
                     label_size, 
                     n = 10000, 
                     from = 0, 
                     to = 10, 
                     by = .01){
  
  d <- tibble(x = rexp(n, rate)) %>%
    ggplot(aes(x = x)) +
    geom_ribbon(aes(ymin = 0,
                    ymax = dexp(x, rate = rate)),
              fill = color_lst[2],
              alpha = 1) +
    stat_pointinterval(aes(y = 0), .width = c(.5, .95)) +
    labs(title = TeX("$\\exp(0.5)$"), # Prior predictive distribution
         x = TeX("Shape ($\\phi$)"),
         y = "Density") + 
    coord_cartesian(xlim = c(0, 10), ylim = (c(0, 0.6))) +
    theme(plot.title = element_text(size = title_size, hjust = 0.5),
          legend.title = element_blank(),
          axis.text = element_text(size = tick_size),
          axis.title = element_text(size = label_size),
          legend.text = element_text(size = tick_size)) 
  
  return(d)
}

```

```{r}

p_exp <- plot_exp(rate = 0.5, 
                  title_size = title, 
                  tick_size = tick,
                  label_size = label)


```

# gamma-Poisson:
* NB(mu, shape/phi): or Gamma-Poisson(mu, shape) #https://bookdown.org/content/4857/monsters-and-mixtures.html
** we infer mu from the data (through linear model).
** we infer shape (phi) from the data (through exponential prior).

As such, we do not place priors directly on these parameters, but we can visualize how the shape 
parameter affects the distribution through running various values through. 

** shape/phi: we get from the gamma (or now exponential) prior.
** mu: we get from our log-likelihood (positive rate?). 

* The NB(mu, phi): 
--> shape is Poisson rate?
--> phi/shape is the gamma prior. (which we show below)


## Maybe it should be gamma, as in: https://bookdown.org/content/4857/monsters-and-mixtures.html

## using dgpois

### specify exact values (rely on intercept = 10 as that is our prior mean for intercept for each group)
### NB: could sample from our prior (Exp(0.5)) instead actually. 

```{r}

### gamma-poisson function ###
plot_dgpois <- function(title_size, 
                        tick_size, 
                        label_size, 
                        shape,
                        intercept){
  
  d <- expand.grid(x = seq(from = 0, to = 50, by = 1),
                   shape = shape) %>%
    mutate(theta = intercept/shape,
           density = dgpois(x, shape = shape, scale = theta)) %>%
  
    ggplot(aes(x = x, y = density, color = as.factor(shape))) + 
    geom_line(size = 2) + 
    labs(title = "Gamma-Poisson", # Prior predictive distribution 
         x = TeX("$c_5$"), # given by exp(\\mu) 
         y = "Frequency") + 
    coord_cartesian(xlim = c(0, 50), ylim = (c(0, 0.25))) +
    theme(plot.title = element_text(size = title_size, hjust = 0.5),
          legend.title = element_text(size = label_size), #lement_blank(),
          axis.text = element_text(size = tick_size),
          axis.title = element_text(size = label_size),
          legend.text = element_text(size = label_size),
          legend.position = c(0.87, 0.87)) +
    scale_color_manual(values = color_lst,
                       name = TeX("Shape ($\\phi$)"))
  
  return(d)
}


```


```{r}

p_gpois <- plot_dgpois(title_size = title,
                       tick_size = tick,
                       label_size = label,
                       shape =  c(0.5, 1, 3),
                       intercept = 10)

```

## gather plot (exp, gamma-pois)

```{r}

p_grid <- plot_grid(p_exp,
                    p_gpois,
                    ncol = 2,
                    labels = c("A", "B"))


```

## save plot

```{r}

ggsave(filename = paste0(outpath, "gamma_poisson.pdf"),
       plot = p_grid,
       height = 4,
       width = 8)

```


## gather plot (exp, gamma)

```{r}

p_grid <- plot_grid(p_exp,
                    p_gamma,
                    ncol = 2,
                    labels = c("A", "B"))

```

```{r}

ggsave(filename = paste0(outpath, "gamma_exp.pdf"),
       plot = p_grid,
       height = 4,
       width = 8)

```

