---
title: "Likelihood Comparison"
author: "Victor M. Poulsen" 
output: html_document
---

# VMP 2022-03-09: should be done now. 

## paths setup


```{r}

wd_code <- "/work/50114/MAG/modeling/replication_fos/code"
inpath <- "/work/50114/MAG/data/modeling/"
infile <- "psychology_replication_matched.csv"
outpath_models <- "/work/50114/MAG/modeling/replication_fos/models/likelihood_comparison/"
outpath_fig <- "/work/50114/MAG/fig/modeling/replication_fos/likelihood_comparison/"

```


# setup 

```{r setup, include=FALSE}
# consider pacman
install.packages("pacman") 
library(pacman)

p_load(tidyverse, brms, ggthemes, bayesplot, cowplot, tidybayes, modelr)

# set up cmdstanr if it is not already present
if (!require('cmdstanr')){
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
library(cmdstanr)
install_cmdstan(cores = 2, overwrite = TRUE)
}

setwd(wd_code)
source("fun_helper.R")
```

## visual setup

consider (a) text-size and (b) width/height

```{r}

# text size: 
# https://statisticsglobe.com/change-font-size-of-ggplot2-plot-in-r-axis-text-main-title-legend

# ggsave: 
# https://ggplot2.tidyverse.org/reference/ggsave.html


color_scheme_set("orange")
theme_set(theme_classic())

theme_set(theme_classic())
title = 18
label = 14
tick = 12 # same as legend

```


# read data and create some variables
## (1) temsize log. 
## (2) match group as id 

```{r}

csv_path <- paste0(inpath, infile)

# check data
d <- read_csv("/work/50114/MAG/data/modeling/psychology_replicat_matched.csv") %>%
  mutate(log_teamsize = log(n_authors), 
         condition_fct = as_factor(condition), 
         id_match = as_factor(match_group),
         id_dct = as_factor(PaperId),
         year_after_2005 = Year - 2005) %>% 
  glimpse()

```

# define model

```{r}

f <- bf(c_5 ~ 0 + condition_fct + condition_fct:log_teamsize + condition_fct:year_after_2005 + (0 + condition_fct | id_match))

```

# get priors 

## zero-inflated poisson: 
b, cor, sd, zi

## negative binomial: 
b, cor, sd, shape 

## zero-inflated negative binomial: 
b, cor, sd, shape, zi 


```{r}

get_prior(
  formula = f, 
  data = d, 
  family = negbinomial()
)

get_prior(
  formula = f,
  data = d,
  family = zero_inflated_poisson(),
)

get_prior(
  formula = f,
  data = d,
  family = zero_inflated_negbinomial()
)

```

# check data

```{r}

d %>% filter(n_authors == 1) %>%
  summarize(mean = mean(c_5))

```


# define priors 

```{r}

## prior negative binomial (also the priors for main model)
p_negbin <- c(prior(exponential(0.5), class = shape), # overdispersion parameter
              prior(normal(log(10), 0.5), class = b, coef = "condition_fctcontrol"),
              prior(normal(log(10), 0.5), class = b, coef = "condition_fctexperiment"),
              prior(normal(0.5, 0.5), class = b, coef = "condition_fctcontrol:log_teamsize"),
              prior(normal(0.5, 0.5), class = b, coef = "condition_fctexperiment:log_teamsize"),
              prior(normal(0, 0.5), class = b, coef = "condition_fctcontrol:year_after_2005"),
              prior(normal(0, 0.5), class = b, coef = "condition_fctexperiment:year_after_2005"),
              prior(exponential(1), class = sd),
              prior(lkj(5), class = cor)) # Solomon Kurz (less flat). 

## prior zero-inflated poisson
p_zip <- c(prior(beta(2, 2), class = zi), # overdispersion parameter
           prior(normal(log(10), 0.5), class = b, coef = "condition_fctcontrol"),
           prior(normal(log(10), 0.5), class = b, coef = "condition_fctexperiment"),
           prior(normal(0.5, 0.5), class = b, coef = "condition_fctcontrol:log_teamsize"),
           prior(normal(0.5, 0.5), class = b, coef = "condition_fctexperiment:log_teamsize"),
           prior(normal(0, 0.5), class = b, coef = "condition_fctcontrol:year_after_2005"),
           prior(normal(0, 0.5), class = b, coef = "condition_fctexperiment:year_after_2005"),
           prior(exponential(1), class = sd),
           prior(lkj(5), class = cor)) # Solomon Kurz (less flat). 

## prior zero-inflated negative binomial
p_zinegbin <- c(prior(exponential(0.5), class = shape), # overdispersion parameter
                prior(beta(2, 2), class = zi),
                prior(normal(log(10), 0.5), class = b, coef = "condition_fctcontrol"),
                prior(normal(log(10), 0.5), class = b, coef = "condition_fctexperiment"),
                prior(normal(0.5, 0.5), coef = "condition_fctcontrol:log_teamsize"),
                prior(normal(0.5, 0.5), coef = "condition_fctexperiment:log_teamsize"),
                prior(normal(0, 0.5), class = b, coef = "condition_fctcontrol:year_after_2005"),
                prior(normal(0, 0.5), class = b, coef = "condition_fctexperiment:year_after_2005"),
                prior(exponential(1), class = sd),
                prior(lkj(5), class = cor)) # Solomon Kurz (less flat). 

```

# sample prior (NB: reproducibility with seed from fun_helper.R)

```{r, include=FALSE}

## sample prior for negative binomial (no divergences)
m_prior_negbin <- fit_model(
  family = negbinomial(), 
  formula = f,
  prior = p_negbin,
  sample_prior = "only",
  file = paste0(outpath_models, "m_prior_negbin")
)

## sample prior for zero-inflated poisson (no divergences)
m_prior_zip <- fit_model(
  family = zero_inflated_poisson(),
  formula = f,
  prior = p_zip,
  sample_prior = "only",
  file = paste0(outpath_models, "m_prior_zip")
)

## sample prior for zero-inflated negative binomial (no divergences)
m_prior_zinegbin <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f,
  prior = p_zinegbin, 
  sample_prior = "only",
  file = paste0(outpath_models, "m_prior_zinegbin")
)

```

# sample posterior 

```{r, include=FALSE}

## sample posterior for negative binomial (174 pareto_k > 0.7 & 136 (8.7%) p_waic > 0.4)
m_post_negbin <- fit_model(
  family = negbinomial(), 
  formula = f,
  prior = p_negbin,
  sample_prior = TRUE,
  file = paste0(outpath_models, "m_post_negbin") 
)

## sample posterior for zero-inflated poisson (1086 pareto_k > 0.7 & 1222 (78.0%) p_waic > 0.4)
m_post_zip <- fit_model(
  family = zero_inflated_poisson(),
  formula = f,
  prior = p_zip,
  sample_prior = TRUE,
  file = paste0(outpath_models, "m_post_zip") 
)

## sample posterior for zero-inflated negative binomial (405 pareto_k > 0.7 & 383 (24.5%) p_waic > 0.4)
m_post_zinegbin <- fit_model(
  family = zero_inflated_negbinomial(),
  formula = f,
  prior = p_zinegbin, 
  sample_prior = TRUE,
  file = paste0(outpath_models, "m_post_zinegbin") 
)

```

# combo plot

```{r}

plot_combo <- function(m_post, filename, tick_size, label_size, height = 11, width = 8){
  
  p <- as_draws_df(m_post) %>%
  rename(#chain = `.chain`,
         `beta[control]` = `b_condition_fctcontrol`,
         `beta[experiment]` = `b_condition_fctexperiment`,
         `beta[control:teamsize]` = `b_condition_fctcontrol:log_teamsize`,
         `beta[experiment:teamsize]` = `b_condition_fctexperiment:log_teamsize`,
         `beta[control:year]` = `b_condition_fctcontrol:year_after_2005`,
         `beta[experiment:year]` = `b_condition_fctexperiment:year_after_2005`) %>%
  mcmc_combo(combo = c("dens_overlay", "trace"),
             pars = vars(contains("beta")),
             facet_args = list(labeller = label_parsed),
             gg_theme = theme(legend.position = "none",
                              strip.text.y = element_text(angle = 0),
                              axis.text = element_text(size = tick_size),
                              axis.title = element_text(size = label_size),
                              strip.background = element_blank()) + facet_text(size=label_size)) 
  
  ggsave(filename = filename, 
         plot = p,
         height = 11,
         width = 8)
  
}

             
```

```{r}

plot_combo(m_post = m_post_negbin,
           filename = paste0(outpath_fig, "negbin_combo.pdf"),
           tick_size = tick,
           label_size = label)

plot_combo(m_post = m_post_zip,
           filename = paste0(outpath_fig, "zip_combo.pdf"),
           tick_size = tick,
           label_size = label)

plot_combo(m_post = m_post_zinegbin,
           filename = paste0(outpath_fig, "zinegbin_combo.pdf"),
           tick_size = tick,
           label_size = label)

```

# pareto

```{r}

pareto_k <- function(d, m, likelihood){
  # d: dataframe 
  # m: model 
  # likelihood: string (for printing only)
  
  d_pareto <- d %>%
    mutate(k = m$criteria$loo$diagnostics$pareto_k) %>%
    filter(k > .7) %>%
    summarize(pareto_k = n()) 

  return(d_pareto)
  
}

```

```{r}

pareto_negbin <- pareto_k(d, m_post_negbin, "pareto_k > 0.7 (negative binomial)") # 195
pareto_zip <- pareto_k(d, m_post_zip, "pareto_k > 0.7 (zero-inflated poisson)") # 1086
pareto_zinegbin <- pareto_k(d, m_post_zinegbin, "pareto_k > 0.7 (zero-inflated negative binomial)") # 447

```

# Rhat + Eff. samples

```{r}

summary_diagnostics <- function(m){ # m = model
  
  # get brmssummary object
  m_summary <- print(m)
  
  # get data
  d_fixed <- as.data.frame(m_summary$fixed) %>% select(Rhat, Bulk_ESS, Tail_ESS)
  d_family <- as.data.frame(m_summary$spec_pars) %>% select(Rhat, Bulk_ESS, Tail_ESS)
  d_group <- as.data.frame(m_summary$random) %>% rename_with(~ gsub("id_match.", "", .)) %>% select(Rhat, Bulk_ESS, Tail_ESS)
  
  # summarize 
  d_summary <- bind_rows(d_fixed, d_family, d_group) %>%
    summarize(max_rhat = max(Rhat),
            min_tail = min(Tail_ESS),
            min_bulk = min(Bulk_ESS))
  
  # return
  return(d_summary)
  
}
```

```{r, include=FALSE}

summary_negbin <- summary_diagnostics(m_post_negbin)
summary_zip <- summary_diagnostics(m_post_zip)
summary_zinegbin <- summary_diagnostics(m_post_zinegbin)

```

# cbind and save 

```{r}

diagnostics_negbin <- cbind(summary_negbin, pareto_negbin)
diagnostics_zip <- cbind(summary_zip, pareto_zip)
diagnostics_zinegbin <- cbind(summary_zinegbin, pareto_zinegbin)

write_csv(diagnostics_negbin, paste0(outpath_fig, "diagnostics_negbin.csv"))
write_csv(diagnostics_zip, paste0(outpath_fig, "diagnostics_zip.csv"))
write_csv(diagnostics_zinegbin, paste0(outpath_fig, "diagnostics_zinegbin.csv"))

```

