---
title: "Untitled"
output: html_document
---

```{r, include=FALSE}

# consider pacman
if (!require("pacman")){
  install.packages("pacman") # repos = "http://cran.r-project.org"
}

library(pacman)
p_load(tidyverse, brms, ggthemes, bayesplot, cowplot, tidybayes, modelr, latex2exp, ggpubr)

# set up cmdstanr if it is not already present
if (!require('cmdstanr')){
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
library(cmdstanr)
install_cmdstan(cores = 2, overwrite = TRUE)
}


```

```{r}

color_scheme_set("orange")
theme_set(theme_classic())

theme_set(theme_classic())
title = 18
label = 14
tick = 12 # same as legend

color_map <- c("Control" = '#1b9e77', 
               "Experiment" = '#d95f02')

categories <- c("Control",
                "Experiment")

clrs = c("#d95f02", "#1b9e77") # a bit hacky.

```

# FOS 

```{r, include=FALSE}

inpath <- "/work/50114/MAG/data/modeling/"
infile <- "psychology_replicat_matched.csv"
outpath <- "/work/50114/MAG/fig/modeling/replication_keyword/results/"
inpath_post <- "/work/50114/MAG/modeling/models/replication_fos/main_model/m_post.rds"
csv_path <- paste0(inpath, infile)

# check data
d_fos <- read_csv(csv_path) %>%
  mutate(log_teamsize = log(n_authors), 
         condition_fct = as_factor(condition), 
         condition_fct = fct_relevel(condition_fct, c("experiment", "control")),
         id_match = as_factor(match_group),
         id_dct = as_factor(PaperId),
         year_after_2005 = Year - 2005) 

```


```{r}

m_fos <- readRDS(paste0(inpath_post))

```


# EVIDENCE RATIO

## mean values of YEAR and TEAMSIZE

```{r}

d_sum <- d_fos %>% summarize(mean_team_log = mean(log_teamsize), # 0.9076
                             mean_team_nat = exp(mean_team_log), # 2.478
                             mean_year = mean(year_after_2005), # 6.244
                             mean_year_nat = mean_year + 2005) # 2011.24

# save
write_csv(d_sum, paste0(outpath, "data_statistics.csv"))

```
## get values out 

```{r}

mean_team_log <- d_sum$mean_team_log
mean_year <- d_sum$mean_year

```


Our group-difference for the INTERCEPT parameter is the expected difference in citations between
replication (experiment) and non-replication (control) studies for a solo-authored paper published
in 2005. A better idea about which group of studies (replication, control) is more cited on average
is obtained when comparing the expected citation difference for mean values of our predictors, 
TEAMSIZE and YEAR. The mean TEAMSIZE in the R_{FOS} data set is 2.48 authors and the mean YEAR 
in our data is 2011.24. When comparing the inferred parameters for replication and control studies
with these values, the expected citation difference is significant (Est: $3.3$, CIs: $1.6, 5.07$, Evid.Ratio $> 1000$, Post.Prob = $1$). All posterior probability is consistent with replication (experiment) studies being cited more
than non-replication (control) studies for mean values of TEAMSIZE and YEAR.

```{r}

# run hypothesis
d_means_int <- hypothesis(m_fos, glue("exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*{mean_team_log} + condition_fctexperiment:year_after_2005*{mean_year}) - exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*{mean_team_log} + condition_fctcontrol:year_after_2005*{mean_year}) > 0"), 
alpha = 0.05)

# gather dataframe 
d_means_int_csv <- d_means_int$hypothesis
write_csv(d_means_int_csv, paste0(outpath, "intercept_mean_values_interaction.csv"))

```

```{r}

# run hypothesis
d_means_control <- hypothesis(m_fos, glue("exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*{mean_team_log} + condition_fctcontrol:year_after_2005*{mean_year}) > 0"), alpha = 0.05)

# gather dataframe
d_means_control_csv <- d_means_control$hypothesis
write_csv(d_means_control_csv, paste0(outpath, "intercept_mean_values_control.csv"))

```

```{r}

# run hypothesis 
d_means_exp <- hypothesis(m_fos, glue("exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*{mean_team_log} + condition_fctexperiment:year_after_2005*{mean_year}) > 0"), alpha = 0.05)

d_means_exp_csv <- d_means_exp$hypothesis
write_csv(d_means_exp_csv, paste0(outpath, "intercept_mean_values_experiment.csv"))

```

# Visualize 

## distribution

```{r}

d_fos_mean <- d_fos %>%
    group_by(condition_fct) %>%
    data_grid(log_teamsize = mean(log_teamsize), # sequence of values
              year_after_2005 = mean(year_after_2005)) %>% # fixed year
    add_epred_draws(m_fos, re_formula = NA) # no random eff. uncertainty

```

```{r}

x_min = 5
if (grepl("fos", outpath, fixed = TRUE)) { # if it is FOS
  x_max = 15
} else {
  x_max = 20 # if it is QUERY
}

```

```{r}

p_dist <- d_fos_mean %>% ggplot(aes(.epred, condition_fct, fill = condition_fct)) + 
  ggdist::stat_halfeye(point_color = NA, .width = 0, height = 1) +  
  geom_boxplot(width = .1, outlier.shape = NA, position = position_nudge(y = -0.1), show.legend=FALSE) + 

  #xlim(0, 300) +
  scale_y_discrete(expand = c(0.12, 0.12)) +
  scale_x_continuous(limits = c(x_min, x_max), expand = c(0.02, 0.02)) + 
  labs(#title = TeX("Outcome Distribution ($c_5$)"),
       y = "Density", 
       x = TeX("$c_5$")) + 
    scale_fill_manual(values = clrs,
                      guide = guide_legend(title = NULL)) + 
  scale_color_manual(values = clrs,
                     guide = guide_legend(title = NULL)) +
  theme(#plot.title = element_text(hjust = 0.5, size = title),
        axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        legend.text = element_text(size = tick),
        legend.position = "none",
        #axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

```

## predictions

```{r}

base_id = 2000
levels <- nrow(d_fos)/2
max_id = base_id + levels - 1

d_pred <- d_fos %>% dplyr::select(condition_fct, log_teamsize, year_after_2005) %>%
  mutate(id_match = as_factor(rep(base_id:max_id, each = 2))) %>% # new ids 
  add_predicted_draws(m_fos, allow_new_levels = T) # no cutoff

```


```{r}
d_pred %>% group_by(condition_fct) %>% 
  summarize(mean_pred = mean(.prediction))
```


```{r}

library(scales) # to access break formatting functions

text_y = 500000
# adding very small number to avoid log(0). 
p_pred <- d_pred %>% group_by(condition_fct, .prediction) %>% 
  # prepare
  summarize(count = n()) %>%
  mutate(prediction = .prediction,
         N = count) %>% 
  # plot 
  ggplot(aes(x = prediction, y = N)) + #, color = condition_fct)) + 
  geom_freqpoly(data = . %>% filter(condition_fct == "control"), 
                stat = 'identity',
                aes(color = categories[1])) + 
  geom_freqpoly(data = . %>% filter(condition_fct == "experiment"), 
                stat = 'identity',
                aes(color = categories[2])) + 
  geom_vline(xintercept = 10) + 
  geom_vline(xintercept = 100) +
  geom_vline(xintercept = 1000) +
  annotate("text", x = 15, y = text_y, label = TeX("$c_5 = 10$"), size = 4) + 
  annotate("text", x = 160, y = text_y, label = TeX("$c_5 = 100$"), size = 4) + 
  annotate("text", x = 1700, y = text_y, label = TeX("$c_5 = 1000$"), size = 4) + 
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  labs(title = "", # "Model Simulated Data",
       x = TeX("$c_5$"),
       y = "N") +
  scale_color_manual(
      breaks = categories, 
      values = color_map,
      guide = guide_legend(title = NULL)) +
  theme(plot.title = element_text(hjust = 0.5, size = title),
        axis.text = element_text(size = tick),
        axis.title = element_text(size = label),
        legend.text = element_text(size = tick),
        legend.position = "bottom") 

```

## gather plots 

```{r}

p_grid <- ggarrange(p_dist, p_pred, ncol = 1, labels = c("A", "B"), common.legend = TRUE, legend = 'bottom')

ggsave(filename = paste0(outpath, "main_intercept_predict.pdf"),
       plot = p_grid,
       width = 8,
       height = 7)

```

## predictors (YEAR)

```{r}

# run hypothesis
d_year_int <- hypothesis(m_fos, "(exp(condition_fctexperiment + condition_fctexperiment:year_after_2005*10) - 
                         exp(condition_fctexperiment)) - (exp(condition_fctcontrol + condition_fctcontrol:year_after_2005*10) - exp(condition_fctcontrol)) > 0", alpha = 0.05)

# gather results
d_year_int <- d_year_int$hypothesis
write_csv(d_year_int, paste0(outpath, "year_10_interaction.csv"))



```

```{r}

# run hypothesis
d_year_exp <- hypothesis(m_fos, "exp(condition_fctexperiment + condition_fctexperiment:year_after_2005*10) - 
                         exp(condition_fctexperiment) > 0", alpha = 0.05)

# gather results
d_year_exp <- d_year_exp$hypothesis
write_csv(d_year_exp, paste0(outpath, "year_10_experiment.csv"))


```

```{r}

# run hypothesis
d_year_cont <- hypothesis(m_fos, "exp(condition_fctcontrol + condition_fctcontrol:year_after_2005*10) - 
                         exp(condition_fctcontrol) > 0", alpha = 0.05)

# gather results 
d_year_cont <- d_year_cont$hypothesis
write_csv(d_year_cont, paste0(outpath, "year_10_control.csv"))


```


## plot 

```{r}

d_team_mean <- d_fos %>%
    group_by(condition_fct) %>%
    data_grid(log_teamsize = seq_range(log_teamsize, n = 101), # sequence of values
              year_after_2005 = mean_year) %>% # fixed year
    add_epred_draws(m_fos, re_formula = NA) # no random eff. uncertainty

```


```{r}

plt_team_nat <- function(draws, d, clrs, xlim, ylim){
  
  p <- draws %>% ggplot(aes(x = exp(log_teamsize), 
                       y = c_5, 
                       color = ordered(condition_fct), 
                       fill = ordered(condition_fct))) +
    #geom_jitter(data = d, alpha = 0.1) + 
    stat_lineribbon(aes(y = .epred), .width = c(.95, .80), alpha = 1/4) +
    labs(title = "", #  "Expected Value (mean actor)",
         x = "TEAMSIZE",
         y = TeX("$c_5$")) +
    theme(plot.title = element_text(hjust = 0.5, size = title),
          axis.text = element_text(size = tick),
          axis.title = element_text(size = label),
          legend.text = element_text(size = tick),
          legend.position = "bottom") +
    scale_color_manual(
      values = clrs,
      guide = guide_legend(title = NULL)) +
    scale_fill_manual(
      values = clrs,
      guide = guide_legend(title = NULL)) +
    ylim(0, ylim) +
    xlim(0, xlim)
  
  return(p)
  
}

d_team_sub <- d_team_mean %>% filter(log_teamsize <= log(20))
y_max = quantile(d_team_sub$.epred, .9999)
p_team_sub <- plt_team_nat(d_team_sub, d_fos, clrs, 20, y_max)

```

```{r}

d_year_mean <- d_fos %>%
    group_by(condition_fct) %>%
    data_grid(log_teamsize = mean_team_log, # sequence of values
              year_after_2005 = seq_range(year_after_2005, n = 101)) %>% # fixed year
    add_epred_draws(m_fos, re_formula = NA) # no random eff. uncertainty

```


```{r}

plt_year <- function(draws, d, clrs, ylim){
  
  p <- draws %>% mutate(actual_year = year_after_2005 + 2005) %>%
    ggplot(aes(x = actual_year, 
                       y = c_5, 
                       color = ordered(condition_fct), 
                       fill = ordered(condition_fct))) +
      #geom_jitter(data = d, alpha = 0.8) + 
      stat_lineribbon(aes(y = .epred), .width = c(.95, .80), alpha = 1/4) +
      labs(title = "", #  "Expected Value (mean actor)",
           x = "YEAR",
           y = TeX("$c_5$")) +
      theme(plot.title = element_text(hjust = 0.5, size = title),
            axis.text = element_text(size = tick),
            axis.title = element_text(size = label),
            legend.text = element_text(size = tick),
            legend.position = "bottom") +
      scale_color_manual(
        values = clrs,
        guide = guide_legend(title = NULL)) +
      scale_fill_manual(
        values = clrs,
        guide = guide_legend(title = NULL)) +
    scale_x_continuous(limits = c(2005, 2015), breaks = c(2005, 2007, 2009, 2011, 2013, 2015)) +
      ylim(0, ylim) 
  
  return(p)
}

ymax = max(d_year_mean$.epred) 
p_year <- plt_year(d_year_mean, d, clrs, ymax)

```



### gather plots

```{r}

p_grid <- ggarrange(p_year, p_team_sub, ncol = 2, labels = c("A", "B"), common.legend = TRUE, legend = 'bottom')

ggsave(filename = paste0(outpath, "main_interaction_effs.pdf"),
       plot = p_grid,
       width = 8,
       height = 5)

```

## predictors (TEAMSIZE)

```{r}

#log(2) # 0.693
#log(5) # 1.609
#log(20) # 2.996

```


```{r}

# run hypothesis
d_team_small_medium <- hypothesis(m_fos, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) - 
                         exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.693) > 0", alpha = 0.05)

# gather results
d_team_small_medium <- d_team_small_medium$hypothesis # Est: 2.23 (0.92, 3.84), Evid = 1599, Post.Prob = 1
write_csv(d_team_small_medium, paste0(outpath, "team_small_medium_exp.csv"))

```

```{r}

# hypothesis
d_team_medium_large <- hypothesis(m_fos, "exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*2.996) - 
                         exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) > 0", alpha = 0.05)

# gather results
d_team_medium_large <- d_team_medium_large$hypothesis # Est: 5.03 (1.65, 9.85), Evid = 1599, Post.Prob = 1
write_csv(d_team_medium_large, paste0(outpath, "team_medium_large_exp.csv"))

```

```{r}

# hypothesis
d_team_small_medium <- hypothesis(m_fos, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.693) > 0", alpha = 0.05)

# gather results
d_team_small_medium <- d_team_small_medium$hypothesis 
write_csv(d_team_small_medium, paste0(outpath, "team_small_medium_cont.csv"))

```

```{r}

# hypothesis
d_team_medium_large <- hypothesis(m_fos, "exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*2.996) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) > 0", alpha = 0.05)

# gather results
d_team_medium_large <- d_team_medium_large$hypothesis 
write_csv(d_team_medium_large, paste0(outpath, "team_medium_large_cont.csv"))

```

```{r}

# hypothesis
d_team_small_medium_int <- hypothesis(m_fos, "(exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*0.693)) - (exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609) - exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*0.693)) > 0", alpha = 0.05)

# gather results
d_team_small_medium_int <- d_team_small_medium_int$hypothesis
write_csv(d_team_small_medium_int, paste0(outpath, "team_small_medium_int.csv"))


```


```{r}

# hypothesis
d_team_medium_large_int <- hypothesis(m_fos, "(exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*2.996) - 
                         exp(condition_fctcontrol + condition_fctcontrol:log_teamsize*1.609)) - (exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*2.996) - exp(condition_fctexperiment + condition_fctexperiment:log_teamsize*1.609)) > 0", alpha = 0.05)

# gather results
d_team_medium_large_int <- d_team_medium_large_int$hypothesis
write_csv(d_team_medium_large_int, paste0(outpath, "team_medium_large_int.csv"))

```
